A Type System for String Sanitation\\Implemented Inside a Python

### 2014-04-16 sanitation-plas14/sanitation-plas14.tex
More progess on Sec. 3

\documentclass[10pt,preprint]{sigplanconf}
\newcommand{\lamAce}{\lambda_{\text{Ace}}}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{verbatimbox}
\input{../att-icfp14/macros-atlam}
\usepackage{alltt}
\renewcommand{\ttdefault}{txtt}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{light-gray}{gray}{0.5}
\usepackage{listings}
\usepackage{wasysym}
    \makeatletter

% \btIfInRange{number}{range list}{TRUE}{FALSE}
%
% Test if int number <number> is element of a (comma separated) list of ranges
% (such as: {1,3-5,7,10-12,14}) and processes <TRUE> or <FALSE> respectively
%
        \newcount\bt@rangea
        \newcount\bt@rangeb

@@

\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\lang}[1]{\Lagr\{#1\}}
\newcommand{\stru}[2]{ {\tt string\_union}(#1,#2)}

@@

\newcommand{\reduces}{ \Rightarrow }
\newcommand{\gvd}{\Gamma \vdash }
\newcommand{\ovd}{\Omega \vdash }

@@

\newcommand{\tstrf}[1]{`#1\textrm'} %??
\newcommand{\strf}[1]{``#1"}

%%
%% Source and Target language definitions.
%%
% Source language terms.
\newcommand{\sisubst}[3]{{\tt subst}[#1](#2,#3)}
\newcommand{\coerce}[2]{ {\tt coerce}[#1](#2)}
\newcommand{\sistr}[1]{{\tt si\_str}[#1]}
\newcommand{\strin}[1]{\sistr{#1}}
\newcommand{\siconcat}[2]{{\tt siconcat}(#1,#2)}

% Source language types.
\newcommand{\stringin}[1]{{\tt string\_in}[#1]}

% target language terms.
\newcommand{\tsubst}[3]{{\tt tsubst}(#1,#2,#3)}
\newcommand{\tcheck}[2]{{\tt tcheck}(#1, #2)}
\renewcommand{\tstr}[1]{{{\tt tstr}[#1]}}
\newcommand{\tconcat}[2]{{\tt tconcat}(#1,#2)}

% Target language types
\newcommand{\str}{{\tt string}}
\newcommand{\regext}[1]{ {\tt rx}[#1] }
\newcommand{\rx}[1]{ {\tt rx}[#1] }

% Meta-theoretic functions
\newcommand{\dosubst}[3]{{\tt dosubst}(#1,#2,#3)}
\newcommand{\lsubst}[3]{{\tt lsubst}(#1,#2,#3)}

% Judgements
\newcommand{\err}{{\tt err}}
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

% Relations
\newcommand{\treduces}{ \Downarrow }
\newcommand{\sreduces}{ \Downarrow }

%%
%% Constrain the size of full-page diagrams and rule lists
%%
%%\newcommand{\pagewidth}{5in}
%%\newcommand{\rulelistwidth}{3in}

% Names of type systems presented in paper
\newcommand{\lcs}{\lambda_{CS}}

\setlength{\grammarindent}{3em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

This section presents the grammar and semantics of $\lcs$.
The semantics are defined in terms of an internal language with at least strings and a regex filter function.
These constraints are captured by the internal term valuations.
The internal language does not necessarily need a regex filter function because
any dynamic conversion is easily definable using a combination of filters and safe
casting.

% Whatever just dropping this in somewhere random.
\begin{thm}
If $S : \strin{r}$ then there exists a T such that $\trden{s} = T$ and either: 
\begin{itemize}
\item (a) $T \treduces \tstr{s}$ and $S \sreduces \strin{s}$, and $s \in lang{r}$.
\item (b) $T \err$ and $S \err$.
\end{itemize}
\end{thm}
\begin{proof}
By induction on the derivation of $\trden{S} = T$. TODO-nrf add case magic.
\begin{itemize}
\item Tr-string. By assumption $\trden{S} = \tstr{s}$. The (a) property follows.
By S-E-Str, $S \sreduces \str{s}$ and by T-E-Str, $\tstr{s} \treduces \tstr{s}$.
By source language type safety, it follows that $s \in \lang{r}$.
\item Tr-Concat.
By inversion of the typing relaiton for the source lnaguage, $S_1 : \strin{r_1}$ and $S_2 : \stringin{S_2}$.
By induction on the left premise, $T_1 \treduces \tstr{s_1}$, $S_1 \sreduces \str{s_1}$ and $s_1 \in \lang{r_1}$
or else $T_1 err$. 
Similarly for $T_2$ and $S_2$.
In the error cases, the propagations rules corresponding to S-E-Concat and T-E-Concat imply that $\tconcat{T_1}{T_2} \err$ and $\siconcat{S_1}{S_2} \err$.
In the non-error case, $\siconcat{S_1}{S_2} \sreduces \strin{s_1 \cdot s_2}$ by S-E-Concat and $\tconcat{T_1}{T_2} \treduces \tstr{s_1 \cdot s_2}$ by T-E-Concat.
Finally, $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$ implies that $s_1 \cdot s_2 \in \lang{r}$ by \ref{lemmaregex}.
\item T-subst.
Suppose $\sisubst{r}{S_1}{S_2} : \strin{s}$ for some $x$. 
By inverison, $S_1 : \strin{r_1}$ and $S_2 : \strin{r_2}$.
Note that $s : \subst{r}{S_1}{S_2} : \stringin{\tsubst{r}{r_1}{r_2}}$ by S-T-Subst.
By induction, $\trden{S_1} = T_1$ and $\trden{S_2} = T_2$.
Either (a) or (b) holds for each.

If (a) holds for both, the $S_1 \sreduces \strin{S_1}$ and $S_2 \sreduces \strin{S_2}$ for $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$.
Therefore, $\sisubst{r}{S_1}{S_2} \sreduces \strin{\dosubst{r,s_1,s_2}}$.
The same fact can be shown in an analogous manner for $T_1$ and $T_2$.
Finally, $s_1 \in \lang{r_1}$ and $s_2 \in \lang{r_2}$, which implies that $\dosubst{r}{s_1}{s_2} \in \lang{\sisubst{r}{s_1}{s_2}}$ by \ref{substhm}.

If (b) holds for either the left or right premise, then $\sisubst{r}{S_1}{S_2} \err$ by the propagation rules corresponding to S-T-subst, 
and $\tsubst{S_1}{S_2} \err$ by the propagation rules corresponding to T-T-Subst.
\end{itemize}
\end{proof}

%
% Begin Grammar 
%

@@

<$\psi$> ::=				\hfill	Source language types					\\
$\stringin{r}$				\hfill Regular expression types \alt

<S> ::= \hfill Source language expressions \\
      $\strin{S}$ \alt
      $\siconcat{S}{S}$ \hfill String concatenation \alt
      $\sisubst{r}{S}{S}$ \alt
      $\coerce{r}{S}$

<$\theta$> ::= \hfill Target language types \\
$\str$ \alt $\regext$

<P> ::= \hfill Internal expressions \\
  $\tstr{P}$ \alt
  $\siconcat{P}{P}$ \alt
  $\tsubst{r}{P}{P}$ \alt
  $\tcheck{r}{P}$
\end{grammar}
\caption{Syntax for the string sanitation fragment of $\lcs$}
\label{fig:lcsSyntax}
\end{figure}

%The $\lcs$ language gives static semantics for common regular expression library
%functions. In this treatment, we include concatenation and filtering.
%The ${\tt filter}$ function removes all instances of a regular expression in a string,
%while concatenation ($+$) concatenates two strings.
%
%\subsection{Typing}
%
%The $\strin[r]$ type is parameterized by regular expressions; if $e:\strin[r]$,
%then $e \in r$. Mapping from an arbitrary $\str$ to a $\strin[r]$ requires
%defining an algorithm -- in terms of filter -- for converting a $\strin[.*]$
%into a $\strin[r]$. The static semantics of the language defines the types of
%operations on regular expressions in terms of well-understood properties about
%regular lanugages; we recall these properties in section 3.
%
%\subsection{Dynamics}
%
%There are two evaluation judgements: $e:T \reduces e'$ and $e:T \treduces i$.
%The $\reduces$ relation is between $\lcs$ expressions, while the $\treduces$
%relation is a mapping from $\lcs$ expressions into internal language expressions
%$i$ such that $i \ival$.
%
%Safety of the evaluation relation depends upon an injective mapping from $\lcs$ types info
%internal language types. This relation, $h$, is defined below.
%
%\subsection{Type Safety}
%
%The type safety proof relies upon some assumptions about the type system and
%dynamics of the internal language, as well as some properties of regular
%languages.
%
%There must exist a translation from $\lcs$ types to the types of the internal
%language. For the remainder of this paper, we call the type translation function
%$h$.
%
%\begin{defn}[Type Translation Function $h$]
%  The type translation function $h : Type \rightarrow IType$ is defined as follows:
%  \begin{itemize}
%    \item $\forall r. h(\strin[r]) = istr$
%    \item $h(\str) = istr$
%  \end{itemize}
%\end{defn}
%
%Additionally, we assume that the internal language contains an implementation of strings,
%together with operations for concatenation and filtering by regular expression.
%
%\begin{defn}[Types of internal values]
%  Let $\tstrf{s}$ range over string literals and $r$ over regular expressions.
%  Internal values are typed as follows:
%\begin{itemize}
%  \item If $e=\tstrf{s}$ then $e:istr$.
%  \item If $e=\ifilter{r}{\tstrf{s}}$ then $e:istr$.
%  \item If $e=\tstrf{s_1}+\tstrf{s_2}$ then $e:istr$.
%\end{itemize}
%\end{defn}
%
%For simplicity, we assume a fixed translation from
%$\lcs$ regular expressions to regular expressions recognizable by the internal
%language's regex library (in practice, a fixed translation is acceptable.)
%To summarize, we assume an internal language containing 
%a string type together with operations for string concatentation and filtering. We
%expect closure over strings for both operations.
%Finally, recall that ${\tt ifilter}$ is only needed for dynamic casts, which may be removed without
%descreasing the expressivity or even usability of the language.
%Finally, the semantics of the filter function are defined in terms of ${\tt rl\_filter}$,
%which is a static version of ${\tt ifilter}$.
%
\subsection{Properties of Regular Languages}

@@

\onecolumn
\newcommand{\sctx}{\Gamma} % Context for external typing
\newcommand{\tctx}{\Omega} % Context for internal typing

@@

\inferline{S-T-include}
{\sctx \vdash \strin{s} : \stringin{r}}
{s \in \lang{r}}

\inferline{S-T-Concat}
{\sctx \vdash \siconcat{S_1}{S_2} : \strin{r_1 \cdot r_2}}
{\sctx \vdash S_1 : \strin{r_1} \\ \sctx \vdash S_2 : \strin{r_2}}

\inferline{S-T-subst}
{\sctx \vdash \sisubst{r}{S_1}{S_2} : r'}
{\sctx \vdash S_1 : \strin{r_1} \\ \sctx \vdash S_2 : \strin{r_2} \\ {\tt lsubst}(r,r_1,r_2) = r'}

\inferline{S-T-coerce}
{\sctx \vdash \coerce{r'}{s} : \stringin{r'}}
{\sctx \vdash S : \stringin{r}}

\end{tabular}
\caption{Typing rules for source langauge terms (S). The metavariable s ranges over
string literals.}
\label{fig:etyping}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c r}
\inferline{S-E-Include}
{\strin{s} \sreduces \strin{s}}
{ \cdot }

\inferline{S-E-Concat}
{\siconcat{S_1}{S_2} \sreduces \strin{s_1 \cdot s_2}} % ???
{S_1 \sreduces \strin{s_1} \\ S_2 \sreduces \strin{s_2}}

\inferline{S-E-Subst}
{\sisubst{r}{S_1}{S_2} \sreduces s}
{S_1 \sreduces \strin{s_1} \\ S_2 \sreduces \strin{s_2} \\ \dosubst{r}{s_1}{s_2} = s}

\inferline{S-E-Coerce-OK}
{\coerce{r}{S} \sreduces \strin{s}}
{S \sreduces \strin{s} \\ x \in \lang{r}}

\inferline{S-E-Coerce-NotOk}
{S \err}
{\coerce{r}{S} \sreduces \strin{s} \\ s \not \in \lang{r}}

\end{tabular}
\caption{Big step semantics for the string fragment of the source language.
There are also unmentioned error propagation rules for for S-E-Concat and S-E-Subst.}
\label{fig:eeval}
\end{center}
\end{figure}

\inferline{P-T-String}
{\tctx \vdash \tstr{s} : \str}
{\cdot}

\inferline{P-T-Concat}
{\tctx \vdash \tconcat{P_1}{P_2} : \str}
{\tctx \vdash P_1 : \str \\ \tctx \vdash P_2 : \str}

\inferline{P-T-Subst}
{\tctx \vdash \tsubst{\rx{r}}{P_1}{P_2} : \str}
{\tctx \vdash P_1 : \str \\ \tctx \vdash P_2 : \str}

\inferline{P-T-Check}
{\tctx \vdash \tcheck{\rx{r}}{s} : \str}
{\tctx \vdash P : \str}

\end{tabular}
\caption{Typing rules for target langauge terms (P). The metavariable s ranges over
string literals.}
\label{fig:etyping}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c r}
\inferline{P-E-Str}
{\tstr{s} \treduces \str{s}}
{ \cdot }

\inferline{P-E-Include}
{\rx{r} \treduces \rx{r}}
{ \cdot }

\inferline{P-E-Concat}
{\tconcat{P_1}{P_2} \treduces \tstr{s_1 \cdot s_2}} % ???
{P_1 \treduces \tstr{s_1} \\ P_2 \treduces \tstr{s_2}}

\inferline{P-E-Subst}
{\tsubst{r}{P_2}{P_3} \treduces \tstr{s}}
{P_1 \treduces \rx{r} \\ P_2 \treduces \tstr{s_2} \\ P_3 \treduces \tstr{s_3} \\ \dosubst{r}{s_2}{s_3} = s}

\inferline{P-E-Check-OK}
{\tcheck{r}{P} \treduces \tstr{s}}
{P \treduces \strin{s} \\ s \in \lang{r}}

\inferline{P-E-Check-NotOk}
{P \err}
{\tcheck{r}{T} \treduces \tstr{s} \\ s \not \in \lang{r}}

\end{tabular}
\caption{Big step semantics for the target lanugage.
There are also unimplemented error propagation rules for for Te-Concat and Te-Subst.}
\label{fig:ieval}
\end{center}
\end{figure}

%
% Begin big-step sematnics for internal terms.
%

\inferline{Tr-string}
{\trden{\strin{s}} = \tstr{s}}
{ \ }

\inferline{Tr-Concat}
{\trden{\siconcat{S_1}{S_2}} = \tconcat{P_1}{P_2}}
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2}

\inferline{Tr-Subst}
{\trden{ \sisubst{r}{S_1}{S_2} } = \tsubst{\rx{r}}{P_1}{P_2} }
{ \trden{S_1} = P_1 \\ \trden{S_2} = P_2 }

\inferline{Tr-Subst}
{\trden{ \coerce{r'}{S} } = P}
{ \trden{s} = P \\ S : \strin{r} \\ \lang{r'} \subseteq \lang{r}}

\inferline{Tr-Coerce-NotOk}
{\trden{\coerce{r'}{S}} = \tcheck{\rx{r'}}{S}}
{\trden{S} = P \\ s : \strin{r} \\ \lang{r'} \not \subseteq \lang{r}}

\end{tabular}
\caption{Translation from source terms (S) to target terms (P).
The translation is type-directed in the Tr-Coerce cases.}
\end{center}
\end{figure}

@@

Note that we never insert an internal check where the type of the string implies
that a check must succeed. Furthermore, expensive calculations (such as language
inclusion or ${\tt dsubst}$, all occur at compile time.
Since sanitation problems are generally expressible with smalle, simple expressions,
we believe that the compile-time overhead is not significant enough to prohibiour target usecase. Informal experimentation with our implementation seems to
support this assertion.

@@

\begin{thm}[Main Soundness Result]
  If $\cdot \vdash e : \strin{r} \sreduces i$ then either $i \treduces \err$ or
  if $i \sreduces v$ for $v$??? then $v \in r$.
\end{thm}

@@

%    \item If $e:T$ and $e:T \treduces i$ then $i : \sigma$ such that $h(T) = \sigma$.

@@

%  \item \textbf{E-String}. By the definition of typing for internal terms, $\tstrf{e} : istr$. It suffices to show that $h(\str) = istr$, which follows from the definition of $h$.
%  \item \textbf{E-String\_in}. By the definition of typing for internal terms, $\tstrf{e} : istr$. It suffices to show that $h(\strin[r]) = istr$, which follows from the definition of $h$ for arbitrary $r$.
%  \item \textbf{E-concatval}. By the definition of typing for internal terms, $\tstrf{e_1} + \tstrf{e_2} : istr$. So it suffices to show that $h(\strin[r_1+r_2]) = istr$, which follows from the definition of $h$ for arbitrary $r_1$, $r_2$.

@@

%  \item \textbf{E-DConvertval}. By the definition of typing for internal terms, $\ifilter(r,\tstrf{e}) : istr$. It suffices to show that $h(\strin[r_1+r_2]) = istr$, which follows from the definition of $h$.

>>>>

\documentclass[10pt,preprint]{sigplanconf}
\newcommand{\lamAce}{\lambda_{\text{Ace}}}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{listings}
\usepackage[usenames,dvipsnames]{color}
\usepackage{mathpartir}
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{verbatimbox}
\input{../att-icfp14/macros-atlam}
\usepackage{alltt}
\renewcommand{\ttdefault}{txtt}
\usepackage{color}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{mauve}{rgb}{0.58,0,0.82}
\definecolor{light-gray}{gray}{0.5}
\usepackage{listings}
\usepackage{wasysym}
    \makeatletter
%
% \btIfInRange{number}{range list}{TRUE}{FALSE}
%
% Test if int number <number> is element of a (comma separated) list of ranges
% (such as: {1,3-5,7,10-12,14}) and processes <TRUE> or <FALSE> respectively
%
        \newcount\bt@rangea
        \newcount\bt@rangeb

@@

\newcommand{\Lagr}{\mathcal{L}}
\newcommand{\strin}{{\tt string\_in}}
\newcommand{\lang}[1]{\Lagr\{#1\}}
\newcommand{\stru}[2]{ {\tt string\_union}(#1,#2)}

\newcommand{\str}{ {\tt string} }

@@

\newcommand{\reduces}{ \Rightarrow }
\newcommand{\ireduces}{ \Downarrow }
\newcommand{\sreduces}{ \leadsto }
\newcommand{\gvd}{\Gamma \vdash }
\newcommand{\ovd}{\Omega \vdash }

@@

\newcommand{\val}{ \ {\tt val} }
\newcommand{\ival}{ \ {\tt ival} }

\newcommand{\istrf}[1]{`#1\textrm'} %??
\newcommand{\strf}[1]{``#1"}

%
% Constrain the size of full-page diagrams and rule lists
%
%\newcommand{\pagewidth}{5in}
%\newcommand{\rulelistwidth}{3in}

%
% Names of type systems presented in paper
%
\newcommand{\lcs}{\lambda_{CS}}

\setlength{\grammarindent}{3em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

This section presents the grammar and semantics of $\lcs$.
The semantics are defined in terms of an internal language with at least strings and a regex filter function.
These constraints are captured by the internal term valuations ($\ival$).
The internal language does not necessarily need a regex filter function because
any dynamic conversion is easily definable using a combination of filters and safe
casting.

%
% Begin Grammar 
%
\newcommand{\ereplace}[3]{{\tt replace}[#1](#2,#3)}
\newcommand{\ecoerce}[2]{{\tt coerce}[#1](#2)}
\newcommand{\estr}[1]{{\tt str\_in[#1]}}
\newcommand{\estrt}[1]{{\tt string\_in[#1]}}
\newcommand{\econcat}[2]{{\tt concat}(#1,#2)}

\newcommand{\ireplace}[3]{{\tt ireplace}(#1,#2,#3)}
\newcommand{\icheck}[2]{{\tt icheck}(#1, #2)}
\newcommand{\istr}[1]{{{\tt istr}[#1]}}
\newcommand{\istrt}{{\tt istrt}}
\newcommand{\iconcat}{\hat{+}}
\newcommand{\regext}{ {\tt regex} }

% Special stuff for the semantics.
\newcommand{\replacefn}[3]{\tt replacement(#1,#2,#3)}
\newcommand{\crl}[3]{{\tt compute\_replacement\_lanugage}(#1,#2,#3)}

@@

<$\psi$> ::=				\hfill	External types					\\
$\estrt{r}$				\hfill Regular expression types \alt

<T> ::= \hfill External expressions \\
      $\estr{s}$ \alt
      $\econcat{e}{e}$ \hfill String concatenation \alt
      $\ereplace{r}{e}{e}$ \alt
      $\ecoerce{r}{e}$

<$\theta$> ::= \hfill Internal types \\
$\istrt$ \alt $\regext$

<S> ::= \hfill Internal expressions \\
  $\istr{s}$ \alt
  $\econcat{i}{i}$ \alt
  $\ireplace{r}{i}{i}$ \alt
  $\icheck{r}{i}$
\end{grammar}
\caption{Syntax for the string sanitation fragment of $\lcs$}
\label{fig:lcsSyntax}
\end{figure}

%The $\lcs$ language gives static semantics for common regular expression library
%functions. In this treatment, we include concatenation and filtering.
%The ${\tt filter}$ function removes all instances of a regular expression in a string,
%while concatenation ($+$) concatenates two strings.
%
%\subsection{Typing}
%
%The $\strin[r]$ type is parameterized by regular expressions; if $e:\strin[r]$,
%then $e \in r$. Mapping from an arbitrary $\str$ to a $\strin[r]$ requires
%defining an algorithm -- in terms of filter -- for converting a $\strin[.*]$
%into a $\strin[r]$. The static semantics of the language defines the types of
%operations on regular expressions in terms of well-understood properties about
%regular lanugages; we recall these properties in section 3.
%
%\subsection{Dynamics}
%
%There are two evaluation judgements: $e:T \reduces e'$ and $e:T \ireduces i$.
%The $\reduces$ relation is between $\lcs$ expressions, while the $\ireduces$
%relation is a mapping from $\lcs$ expressions into internal language expressions
%$i$ such that $i \ival$.
%
%Safety of the evaluation relation depends upon an injective mapping from $\lcs$ types info
%internal language types. This relation, $h$, is defined below.
%
%\subsection{Type Safety}
%
%The type safety proof relies upon some assumptions about the type system and
%dynamics of the internal language, as well as some properties of regular
%languages.
%
%There must exist a translation from $\lcs$ types to the types of the internal
%language. For the remainder of this paper, we call the type translation function
%$h$.
%
%\begin{defn}[Type Translation Function $h$]
%  The type translation function $h : Type \rightarrow IType$ is defined as follows:
%  \begin{itemize}
%    \item $\forall r. h(\strin[r]) = istr$
%    \item $h(\str) = istr$
%  \end{itemize}
%\end{defn}
%
%Additionally, we assume that the internal language contains an implementation of strings,
%together with operations for concatenation and filtering by regular expression.
%
%\begin{defn}[Types of internal values]
%  Let $\istrf{s}$ range over string literals and $r$ over regular expressions.
%  Internal values are typed as follows:
%\begin{itemize}
%  \item If $e=\istrf{s}$ then $e:istr$.
%  \item If $e=\ifilter{r}{\istrf{s}}$ then $e:istr$.
%  \item If $e=\istrf{s_1}+\istrf{s_2}$ then $e:istr$.
%\end{itemize}
%\end{defn}
%
%For simplicity, we assume a fixed translation from
%$\lcs$ regular expressions to regular expressions recognizable by the internal
%language's regex library (in practice, a fixed translation is acceptable.)
%To summarize, we assume an internal language containing 
%a string type together with operations for string concatentation and filtering. We
%expect closure over strings for both operations.
%Finally, recall that ${\tt ifilter}$ is only needed for dynamic casts, which may be removed without
%descreasing the expressivity or even usability of the language.
%Finally, the semantics of the filter function are defined in terms of ${\tt rl\_filter}$,
%which is a static version of ${\tt ifilter}$.
%
\subsection{Properties of Regular Languages}

@@

\onecolumn
\newcommand{\ectx}{?} % Context for external typing
\newcommand{\ictx}{?} % Context for internal typing

@@

\inferline{Te-include}
{\ectx \vdash \estr{s} : \estrt{r}}
{s \in \lang{r}???}

\inferline{Te-Concat}
{\ectx \vdash \econcat{S_1}{S_2} : \estrt{r_1 \cdot r_2}}
{\ectx \vdash S_1 : \estrt{r_1} \\ \ectx \vdash S_2 : \estrt{r_2}}

\inferline{Te-subst}
{\ectx \vdash \ereplace{r}{S_1}{S_2} : r}
{\ectx \vdash S_1 : \estrt{r_1} \\ \ectx \vdash S_2 : \estrt{r_2} \\ r = {\tt lsubst}(r,r_1,r_2)}

\inferline{Te-coerce}
{\ectx \vdash \ecoerce{r'}{s} : \estrt{r'}}
{\ectx \vdash S : \estrt{r}}

\end{tabular}
\caption{Typing rules for $\lcs$. Err cases?}
\label{fig:etyping}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c r}
\inferline{Te-Include}
{\estr{s} \val}
{ \cdot }

\inferline{Te-Concat}
{\econcat{S_1}{S_2} \ereduces \estr{s_1 \cdot s_2}} % ???
{S_1 \Downarrow \estr{s_1} \\ S_2 \Downarrow \estr{s_2}}

\inferline{Te-Subst}
{\ereplace{r}{S_1}{S_2} \ereduces s}
{S_1 \Downarrow \estr{s_1} \\ S_2 \Downarrow \estr{s_2} \\ s = {\tt substitution}(r,s_1,s_2)}

\inferline{Te-Coerce-OK}
{\ecoerce{r}{s} \ereduces \estr{x}}
{s \ereduces \estr{x} \\ x \in \lang{r}}

\inferline{Te-Coerce-NotOk}
{\ecoerce{r}{s} \ereduces \estr{x} \\ x \not \in \lang{r}}
{s \ereduces \estr{x} \\ x \not \in \lang{r}}
\end{tabular}
\caption{Operational semantics for the external language $\lcs$.
There are also unmentioned error propagation rules for for Te-Concat and Te-Subst.}
\label{fig:eeval}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\begin{tabular}{c r}
\inferline{Ti-Include}
{\istr{s} \val}
{ \cdot }

\inferline{Ti-Include}
{{\tt regex}[r] \val}
{ \cdot }

\inferline{Ti-Concat}
{\iconcat{S_1}{S_2} \ireduces \istr{s_1 \cdot s_2}} % ???
{S_1 \ireduces \istr{s_1} \\ S_2 \ireduces \istr{s_2}}

\inferline{Te-Subst}
{\ireplace{r}{S_2}{S_3} \ireduces {\tt substitution(r,s_2,s_3)}}
{T_1 \ireduces {\tt regex}[r] \\ T_2 \ireduces \istr{s_2} \\ T_3 \ireduces \istr{s_3} \\ }

\inferline{Te-Ccheck-OK}
{\icheck{r}{s} \ireduces \istr{x}}
{s \ireduces \estr{x} \\ x \in \lang{r}}

\inferline{Te-Check-NotOk}
{\icheck{r}{s} \ireduces \istr{x} \\ x \not \in \lang{r}}
{s \ireduces \istr{x} \\ x \not \in \lang{r}}
\end{tabular}
\caption{Operational semantics for the internal language of $\lcs$.
There are also unmentioned error propagation rules for for Te-Concat and Te-Subst.}
\label{fig:ieval}
\end{center}
\end{figure}

%
%\begin{figure}
%\begin{center}
%\begin{tabular}{c r}
%
%\inferline{T-Abs}
%{\gvd \lambda x : \tau .e : \tau \rightarrow \tau' \sreduces \lambda x : \sigma.i}
%{\Gamma, x : \tau \vdash e : \tau' \sreduces i \\ \sigma \equiv \trep(\tau)}
%
%\inferline{T-App}
%{\Gamma \vdash e_1 e_2 \sreduces i_1 i_2}
%{\Gamma \vdash e_1 : \tau \rightarrow \tau' \sreduces i_1 \\ \Gamma \vdash e_2 : \tau \sreduces i_2}
%
%\inferline{T-Replace}
%{\gvd \ereplace{r}{e_1}{e_2} : \estrt{r'} \sreduces \ireplace{r}{i_1}{i_2}}
%{\gvd e_1 : \estrt{r_1} \ireduces i_1 \\ \gvd e_2 : \estrt{r_2} \ireduces i_2 \\ r' \equiv \crl{r}{r_2}{r_1}}
%
%\inferline{T-Check-Ok}
%{\gvd \echeck{r'}{e} : \estrt{r'} \sreduces i}
%{\gvd e : \estrt{r} \sreduces{i} \\ \lang{r'} \subseteq \lang{r}}
%
%\inferline{T-Check-NotOk}
%{\gvd \echeck{r'}{e} : \estrt{r'} \sreduces \icheck{r}{i}}
%{\gvd e : \estrt{r} \sreduces i \\ \lang{r'} \not \subseteq \lang{r}}
%  
%%\inferline{E-dconvert (optional) }
%%{\dconvert{\strin[r']}{e} : \strin[r'] \ireduces \ifilter{r'}{\istrf{e}}}
%%{e \val }
%
%\end{tabular}
%\caption{Type-directed translation to internal terms for $\lcs$}
%\label{fig:eeval}
%\end{center}
%\end{figure}

%
% Begin typing relation for internal terms.
%
\begin{figure}
\begin{center}
\begin{tabular}{c r}
\inferline{Ti-Var}
{\ovd x : \sigma}
{x : \sigma \in \Omega}

% Add in pedantic rules as well? If there's space.
\inferline{Ti-Abs}
{\ovd \lambda x : \sigma . i : \sigma \rightarrow \sigma'}
{\Omega, x : \sigma \vdash i : \sigma'}

\inferline{Ti-App}
{i_1 i_2 : \sigma'}
{\ovd i_1 : \sigma \rightarrow \sigma' \\ \ovd i_2 : \sigma}

\inferline{Ti-Str}
{\ovd \istr{s} : \istrt }
{\cdot}

\inferline{Ti-Replace}
{\ovd \ireplace{r}{i_1}{i_2} : \istrt}
{\ovd i_1 : \istrt \\ \ovd i_2 : \istrt}

\inferline{Ti-Check}
{\ovd \icheck{r}{i_1}{i_2} : \istrt}
{\ovd i_1 : \istrt \\ \ovd i_2 : \istrt}

\inferline{Ti-Concat}
{\ovd i_1 \iconcat i_2 : \istrt}
{\ovd i_1 : \istrt \\ \ovd i_2 : \istrt}

\end{tabular}
\caption{Typing relation on internal terms.}\label{fig:itypes}
\end{center}
\end{figure}

%
% Begin big-step sematnics for internal terms.
%
\newcommand{\err}{{\tt err}}
\begin{figure}
\begin{center}
\begin{tabular}{c r}

  \inferline{Ei-strval}
  {\istr{s} \ival}
  { \cdot }

  \inferline{Ei-Replace}
  {\ireplace{r}{i_1}{i_2} \ireduces i'}
  { i \ireduces \replacefn{r}{i_1}{i_2} }

  \inferline{Ei-CheckOk}
  {\icheck{r}{i} \ireduces i}
  {i \in \lang{r}}

  \inferline{Ei-Check-NotOk}
  {\icheck{r}{i} \ireduces \err}
  {i \not \in \lang{r}}

  \inferline{Ei-Concat}
  { i_1 \iconcat i_2 \ireduces \istr{i_1 \iconcat i_2} }
  { \cdot }

\end{tabular}
\caption{Big-step semantics for internal language; standard rules for the STL are omitted for brevity. The meta-variable $s$ ranges over string literals.}\label{fig:ieval}
\end{center}
\end{figure}

@@

Note that we never insert an internal check where the type of the string implies
that a check must succeed. Furthermore, expensive calculations (such as language
inclusion or ${\tt computer\_replacement\_language}$, all occur at compile time.
Since sanitation problems are generally expressible with smalle, simple expressions,
we believe that the compile-time overhead is not significant enough to prohibit
our target usecase. Informal experimentation with our implementation seems to
support this assertion.

@@

\begin{thm}[Main Soundness Result]
  If $\cdot \vdash e : \estrt{r} \sreduces i$ then either $i \ireduces \err$ or
  if $i \sreduces v$ for $v \ival$ then $v \in r$.
\end{thm}

@@

%    \item If $e:T$ and $e:T \ireduces i$ then $i : \sigma$ such that $h(T) = \sigma$.

@@

%  \item \textbf{E-String}. By the definition of typing for internal terms, $\istrf{e} : istr$. It suffices to show that $h(\str) = istr$, which follows from the definition of $h$.
%  \item \textbf{E-String\_in}. By the definition of typing for internal terms, $\istrf{e} : istr$. It suffices to show that $h(\strin[r]) = istr$, which follows from the definition of $h$ for arbitrary $r$.
%  \item \textbf{E-concatval}. By the definition of typing for internal terms, $\istrf{e_1} + \istrf{e_2} : istr$. So it suffices to show that $h(\strin[r_1+r_2]) = istr$, which follows from the definition of $h$ for arbitrary $r_1$, $r_2$.

@@

%  \item \textbf{E-DConvertval}. By the definition of typing for internal terms, $\ifilter(r,\istrf{e}) : istr$. It suffices to show that $h(\strin[r_1+r_2]) = istr$, which follows from the definition of $h$.
